Optimal updating magnitude in adaptive flat-distribution-sampling simulations

### 2017-09-13 doc/invt.tex
allow checkeq.py to check duplicated equation labels

The bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for the average growth.\cite{
dama2014}
%
In the second deduction, we define
the bin average as $\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$,
and the deviation of the bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  .
\label{eq:x_def}
\end{equation}
%
\note{Alternatively,
  we may define $v_{*i}(t)$ such that
  $\rho_i \, e^{-v_{*i}(t)}$ is a normalized distribution,
  i.e.
  $$
  v_{*i}(t) \to -\ln\left(
    \frac{ (p_i/\rho_i) \, e^{ -u_i(t) } }
    { \sum_{j=1}^n p_j \, e^{ -u_j(t) } }
  \right).
  $$
  We can readily show that the two definitions are equivalent
  for small $v_{*i}$.}%
%
Below we will use the notations ``$\bav{}$'' and ``$_{*i}$''
for other quantities.
%
The total error is then given by
%
\begin{equation}
  \Err(t)
  =
  \sum_{i = 1}^n \rho_i \,
  \left\langle v_{*i}^2(t) \right\rangle
  .
\label{eq:error_def}
\end{equation}

@@

We will approximate the histogram fluctuation part, $\xi_i(t)$,
%and hence its deviation from the bin average, $\xi_{*i}(t)$,
as white noise such that
\begin{equation}
  \langle \xi_{*i}(t) \, \xi_{*i}(t') \rangle
  = G_i \, \delta(t - t')
  ,
  \notag
  %\label{eq:G_def}
\end{equation}
for some constant $G_i$,
and $\delta(t)$ is Dirac's delta function.
%
We expect the approximation to hold
for simulations much longer than the autocorrelation time.

From Eqs.
\eqref{eq:vt_diffeq} and \eqref{eq:sh_ave},
we get % a set of decoupled equations
%
\begin{equation}
  \dot v_{*i}(t)
  %=
  %\alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \xi_{*i}(t) \right]
  \approx
  -\alpha(t) \, \left[ v_{*i}(t) - \xi_{*i}(t) \right]
  .
  %\notag
  \label{eq:dxdt_singlebin}
\end{equation}
%
Without the random part, $\xi_{*i}(t)$,
the deviation of the bias potential, $v_{*i}(t)$,
would tend to $0$ at long times, reaching convergence.
%
The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
%
\begin{equation}
  v_{*i}(T)
  =
  v_{*i}(0) \, e^{-q(T)}
  +
  \int_0^T
    \dot{\theta}\bigl( q(t) \bigr) \, \xi_{*i}(t) \, dt,
  \label{eq:xt_solution}
\end{equation}
%
where
%
$q(t) \equiv \int_0^t \alpha(t') \, dt'$,
%
and
%
\begin{align}
  \theta(q')
  &\equiv
  e^{q' - q(T)}.
  \label{eq:theta_def}
\end{align}

@@

From Eqs. \eqref{eq:error_def} and \eqref{eq:xt_solution},
we can write the total error at the end of period $T$
as a sum of a residual error, $E_R(T)$ (from the decay of the initial error)
and an asymptotic error, $E_A(T)$:
%
\begin{align}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T)
  ,
  \label{eq:error_tot}
\end{align}
%
where
\begin{align}
  \Err_R(T)
  &= \Err(0) \, e^{-2 \, q(T)}
  ,
  \label{eq:ER_sbin}
  \\
  \Err_A(T)
  &=
  \sum_{i=1}^n \rho_i
    \int_0^T\!\!\!\int_0^T
    \!\!
    \langle
      \xi_{*i}(t) \xi_{*i}(t')
    \rangle
    \dot \theta\bigl( q(t) \bigr)
    \dot \theta\bigl( q(t') \bigr)
    \, dt dt'
  \notag \\
  &= \Gamma \, \int_0^T \dot \theta^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:EA_sbin}
\end{align}
with
%$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
\approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt$.
%\begin{equation}
%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
%  =
%  \sum_{i=1}^n \rho_i \int_0^T
%    \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt
%  .
%  \notag%\label{eq:Gamma_from_G}
%\end{equation}

The error depends implicitly on the curve, $\alpha(t)$,
or, equivalently, the curve, $q(t)$.
%via $v_{*i}(T)$.
%
%If $\alpha(t)$ is a constant, $a_0$,
%so that $q(t) = a_0 \, t$,
%then after a long period,
%the error approach an equilibrium value
%%
%\begin{equation}
%  \Err_\mathrm{equil.}
%  = \Err_A = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
%  = \frac{1}{2} a_0 \Gamma
%  .
%  \label{eq:error_equil_sbin}
%\end{equation}
%
We wish to find the $q(t)$
that minimizes this error.
%
It is convenient to
vary the curve, $q(t)$ ($0 < t < T$),
under a fixed value of $q(T)$,
since it allows us to focus on
the asymptotic error, $E_A(T)$.
%
The contribution of the residual error
will be considered later
by adjusting the value of $q(T)$.
%
Note that the asymptotic error
has the form of an action of a free particle
at $\theta\bigl( q(t) \bigr)$,
with the endpoints fixed at
$\theta\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$\theta\bigl( q(T) \bigr) = 1$.
%
Thus, the velocity should be a constant
%Since the residual error
%is fixed during the process,
%the asymptotic error,
%resembling the action of a free particle,
%demands a constant velocity
%
\begin{equation}
  \frac{d}{dt} \theta\bigl( q(t) \bigr) = c
  .
\label{eq:dthetadt_const}
\end{equation}
%
Then, by using Eq. \eqref{eq:theta_def}, we have
%
\begin{equation}
  e^{ q(t) - q(T) }
  =
  c \, (t + t_0)
  =
  \frac{ t + t_0 }
       { T + t_0 }
  ,
\label{eq:expqt}
\end{equation}
%
where $t_0$ and $c$ are two constants,
and the latter has been determined
from the $t = T$ case
in the second step.
%
Taking the logarithm,
and differentiating both sides
with respect to $t$,
we get the inverse-time schedule
%
\begin{equation}
  \alpha(t) = \frac{ 1 }{ t + t_0 }
  ,
\label{eq:alpha_invt1}
\end{equation}
%
with Eq. \eqref{eq:alpha_invt} being
the special case of $t_0 = 0$.
%

We now determine the optimal value of $q(T)$,
which was previously held fixed.
%
This is equivalent to choosing a value of $t_0$
because $q(T) = \ln\bigl(1 + \frac{T}{t_0}\bigr)$
from the $t = 0$ case of Eq. \eqref{eq:expqt}.
%
Using Eq. \eqref{eq:alpha_invt1} in
Eqs. \eqref{eq:ER_sbin} and \eqref{eq:EA_sbin}
yields
\begin{align}
  \Err_R(T)
  &= \frac{ \Err(0) \, t_0^2 } { (T + t_0)^2 }
  ,
  \notag
  %\label{eq:ER_sbin1}
  \\
  \Err_A(T)
  &= \frac{ \Gamma \, T } { (T + t_0)^2 }
  ,
  \label{eq:EA_sbin1}
\end{align}
%
and the total has a minimum of
\begin{equation}
  \Err(T)
  =
  \frac{ \Gamma } { T + t_0 }
  ,
  \label{eq:Emin_sbin}
\end{equation}
%
reachable at $t_0 = \Gamma /\Err(0)$.
%
%If we assume that at $t = 0$,
%the system has just completed a long simulation
%under a constant updating magnitude, $a_0$,
%then the initial error can be computed from Eq. \eqref{eq:EA_sbin}
%as
%\begin{equation}
%  \Err(0)
%  = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
%  = \frac{1}{2} a_0 \Gamma
%  ,
%  \label{eq:error_equil_sbin}
%\end{equation}
%and
%%
%\begin{equation}
%  t_0 = \frac{ 2 } { a_0 } = \frac{1}{ \alpha(t = 0) }
%  .
%  \label{eq:t0_sbin}
%\end{equation}
%
%The minimum error can be related to
%the expected histogram fluctuation\cite{zhou2005, zhou2008}
%(cf. Appendix \ref{sec:hfluc}).

@@

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \notag
  %\label{eq:w_balance}
\end{equation}
%
We will label this eigenvector by $k = 0$,
and
%
\begin{equation}
  \phi_{0i} = \rho_i,
\label{eq:eigenmode0}
\end{equation}
%
with $\lambda_0 = 1$.
%
Clearly, Eq. \eqref{eq:eigenmode0}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 0$,
while the orthogonality demands
%
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \delta_{k0}
  .
\label{eq:ortho0}
\end{equation}

@@

Equation \eqref{eq:vt_diffeq_mbin} shows that
in the absence of the fluctuating term, $\tilde \xi_k(t)$,
the deviation of the bias potential would tend to zero
at a rate of $\alpha(t) \, \lambda_k$.
%
Thus, the updating scheme is convergent
only if no eigenvalue, $\lambda_k$, is negative
for $k > 0$.
%
%With a set of heterogeneous eigenvalues,
%the errors of the eigenmodes are reduced at different rates.
%Thus, the optimization problem of the final error
%becomes more complicated and it often
%involves a compromise among eigenmodes.
%
The formal solution is given by
\begin{equation}
  \tilde v_k(T)
  =
  \tilde v_k(0) \, e^{-\lambda_k \, q(T)}
  +
  \int_0^T
    \dot{\theta}_k\bigl( q(t) \bigr) \, \tilde\xi_k(t) \, dt
  ,
  \label{eq:xt_solution}
\end{equation}
where
\begin{equation}
  \theta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  .
  \label{eq:uk_def}
\end{equation}

@@

By Eq. \eqref{eq:parseval},
we can rewrite the error defined in Eq. \eqref{eq:error_def} as
a sum of the errors in individual modes\footnote{If
the error weighting distribution differs from $\rho_i$
in Eq. \eqref{eq:error_def},
the error would contain quadratic cross terms among different eigenmodes.
%
While the solution for the optimal schedule retains the same form,
the mass function defined in Eq. \eqref{eq:mass_func}
will include the corresponding cross terms.
}
%$\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
%
\begin{align}
  \Err(T)
  =
  \sum_{k = 1}^{n - 1}
    \left\langle
      {\tilde v}_k^2(T)
    \right\rangle
  .
  \notag
\end{align}
%
As in Eq. \eqref{eq:error_tot},
we can decompose the error as the sum
of the residual error and the asymptotic error, with
\begin{align}
  \Err_R(T)
  &=
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) }
  ,
  \label{eq:error_res}
  \\
  \Err_A(T)
  &=
  \sum_{k=1}^{n-1}
  \int_0^T\!\!\!\int_0^T
  \!\!
  \langle
    \tilde\xi_k(t)
    \tilde\xi_k(t')
  \rangle
  \theta_k\bigl(q(t)\bigr)
  \theta_k\bigl(q(t')\bigr)
  \, dtdt'
  \notag \\
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot \theta_k^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:error_asym}
\end{align}
%
where
%
%$\Gamma_k$ is computed from
%the autocorrelation function
%of the histogram fluctuation,
%$\bigl\langle {\tilde \xi}_k(0)
%\, {\tilde \xi}_k(t) \bigr\rangle$,
%as:
%
\begin{equation}
  \Gamma_k
  \approx
  \int_{-\infty}^\infty
  \bigl\langle
    \tilde\xi_k(t) \, \tilde\xi_k(0)
  \bigr\rangle
  \, dt
  \to
  \sum_{t = -\infty}^\infty
  \bigl\langle
    \tilde\xi_k(t) \, \tilde\xi_k(0)
  \bigr\rangle
  ,
  \label{eq:Gamma_sum}
\end{equation}
%
with the second expression being
the counterpart in the discrete-time setting.
%
Equation \eqref{eq:Gamma_sum} represents
a first approximation for modeling
a general fluctuation, ${\tilde \xi}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
%
Note that by Eq. \eqref{eq:parseval}, we have
$\sum_{k=1}^{n-1} \bigl\langle {\tilde \xi}_k(t) \, {\tilde \xi}_k(0) \bigr\rangle
= \sum_{i=1}^n \rho_i \bigl\langle \xi_{*i}(t) \, \xi_{*i}(0) \bigr\rangle,$
and summing $t$ from $-\infty$ to $\infty$ yields,
\begin{equation}
  \sum_{k=1}^{n-1} \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma
  .
  \label{eq:Gammak_sum}
\end{equation}
This shows the equivalence of
Eq. \eqref{eq:EA_sbin1} and \eqref{eq:error_asym}
in the case of the single-bin updating scheme, with $\lambda_k \equiv 1$.
%In Appendix \ref{sec:Gamma_measure},
%we discuss a way of estimating $\Gamma_k$
%from an FDS simulation under a constant magnitude.
%

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
To model the initial error for Eq. \eqref{eq:error_res},
we will assume that at $t = 0$
the system has undergone a preliminary FDS simulation
under a constant updating magnitude, $a_0$,
for $T_0$ steps.
%
If the initial bias potential
before the preliminary run is zero,
we have, from Eq. \eqref{eq:xt_solution} [with $T\to T_0, q(t) \to a_0 \, t$],
%
\begin{align}
  \left\langle {\tilde v}_k(0) \right\rangle
  &=
  \tilde u_k \, e^{-\lambda_k \, a_0 \, T_0}
  ,
  \label{eq:xt_eql}
  \\
  %\operatorname{var} \tilde v_k(0)
  %&=
  %\frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
  %\left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
  %,
  %\label{eq:xtvar}
  %\\
  \left\langle \tilde v_k^2(0) \right\rangle
  &=
  %\operatorname{var} \tilde v_k(0) + \zeta_k^2
  %=
  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
  +
  \epsilon_k \, e^{-2 \, \lambda_k \, a_0 \, T_0}
  ,
  \label{eq:xt2_eql}
\end{align}
%
where
$\epsilon_k \equiv \left(\tilde u_k^2 - \frac{1}{2} \Gamma_k \, \lambda_k \, a_0\right) \, e^{-2\,\lambda_k a_0 T_0}$.

@@

The results for the $L = 64$ case
after $10^5 L^2$ steps are shown
in Fig. \ref{fig:pt_hist}.
%
The Gaussian histograms were evenly spaced with similar widths as intended,
%and the overall energy histogram was roughly flat,
as shown in Fig. \ref{fig:pt_hist}(a).
%
We computed the density of states by WHAM,
and found the critical temperature to be $T_c = 0.70163$,
in agreement with a previous study\cite{wang2001pre}.
%
Since umbrellas were relatively narrow,
the logarithmic density of states, $\ln g(E)$,
was roughly captured
by $c_0^{(s)}$ as a function of $E_c^{(s)}$,
as shown Fig. \ref{fig:pt_hist}(b).
%
As discussed in Sec. \ref{sec:age},
$c_1^{(s)}/\sigma_E$ roughly represents
the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
and it manifested
the backbending behavior\cite{kim2006, *kim2007, kim2010}
typical in a first-order phase transition,
as shown in Fig. \ref{fig:pt_hist}(c).
%
The coefficient, $\sqrt 2 \, c_2^{(s)}$,
was indeed above $1.0$
in the unstable region,
where $\beta(E)$ is increasing or, equivalently,
$\ln g(E)$ is locally convex,
as shown in Fig. \ref{fig:pt_hist}(d).

>>>>

The bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for the average growth.\cite{
dama2014}
%
In the second deduction, we define
the bin average as $\bav{v}(t) \equiv \sum_{i=1}^n \rho_i \, v_i(t)$,
and the deviation of the bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  .
\label{eq:x_def}
\end{equation}
%
\note{Alternatively,
  we may define $v_{*i}(t)$ such that
  $\rho_i \, e^{-v_{*i}(t)}$ is a normalized distribution,
  i.e.
  $$
  v_{*i}(t) \to -\ln\left(
    \frac{ (p_i/\rho_i) \, e^{ -u_i(t) } }
    { \sum_{j=1}^n p_j \, e^{ -u_j(t) } }
  \right).
  $$
  We can readily show that the two definitions are equivalent
  for small $v_{*i}$.}%
%
Below we will use the notations ``$\bav{}$'' and ``$_{*i}$''
for other variables.
%
The total error is then given by
%
\begin{equation}
  \Err(t)
  =
  \sum_{i = 1}^n \rho_i \,
  \left\langle v_{*i}^2(t) \right\rangle
  .
\label{eq:error_def}
\end{equation}

@@

We will approximate the histogram fluctuation part, $\xi_{*i}(t)$,
%and hence its deviation from the bin average, $\xi_{*i}(t)$,
as white noise such that
%
\begin{equation}
  \langle \xi_{*i}(t) \, \xi_{*i}(t') \rangle
  = G_i \, \delta(t - t')
  ,
  \label{eq:Gi_def}
\end{equation}
%
for some constant $G_i$,
and $\delta(t)$ is Dirac's delta function.
%
We expect the approximation to hold
for simulations much longer than the autocorrelation time.

From Eqs.
\eqref{eq:vt_diffeq} and \eqref{eq:sh_ave},
we get % a set of decoupled equations
%
\begin{equation}
  \dot v_{*i}(t)
  %=
  %\alpha(t) \, \left[ \langle f_{*i}(t) \rangle + \xi_{*i}(t) \right]
  \approx
  -\alpha(t) \, \left[ v_{*i}(t) - \xi_{*i}(t) \right]
  .
  %\notag
  \label{eq:dxdt_singlebin}
\end{equation}
%
Without the random part, $\xi_{*i}(t)$,
the deviation of the bias potential, $v_{*i}(t)$,
would tend to $0$ at long times, reaching convergence.
%
The formal solution of Eq. \eqref{eq:dxdt_singlebin} is
%
\begin{equation}
  v_{*i}(T)
  =
  v_{*i}(0) \, e^{-q(T)}
  +
  \int_0^T
    \dot{\theta}\bigl( q(t) \bigr) \, \xi_{*i}(t) \, dt,
  \label{eq:vi_solution}
\end{equation}
%
where
%
$q(t) \equiv \int_0^t \alpha(t') \, dt'$,
%
and
%
\begin{align}
  \theta(q')
  &\equiv
  e^{q' - q(T)}.
  \label{eq:theta_def}
\end{align}

@@

From Eqs. \eqref{eq:error_def},
\eqref{eq:vi_solution}, and
\eqref{eq:Gi_def},
we can write the total error at the end of period $T$
as a sum of a residual error, $E_R(T)$ (from the decay of the initial error)
and an asymptotic error, $E_A(T)$:
%
\begin{align}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T)
  ,
  \label{eq:error_tot}
\end{align}
%
where
\begin{align}
  \Err_R(T)
  &= \Err(0) \, e^{-2 \, q(T)}
  ,
  \label{eq:ER_sbin}
  \\
  \Err_A(T)
  &=
  \sum_{i=1}^n \rho_i
    \int_0^T\!\!\!\int_0^T
    \!\!
    \langle
      \xi_{*i}(t) \xi_{*i}(t')
    \rangle
    \dot \theta\bigl( q(t) \bigr)
    \dot \theta\bigl( q(t') \bigr)
    \, dt dt'
  \notag \\
  &= \Gamma \, \int_0^T \dot \theta^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:EA_sbin}
\end{align}
with
%$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i$.
$\Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
\approx \sum_{i=1}^n \rho_i \int_{-\infty}^{\infty} \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt$.
%\begin{equation}
%  \Gamma \equiv \sum_{i=1}^n \rho_i \, G_i
%  =
%  \sum_{i=1}^n \rho_i \int_0^T
%    \langle \xi_{*i}(t) \xi_{*i}(0) \rangle \, dt
%  .
%  \notag%\label{eq:Gamma_from_G}
%\end{equation}

The error depends implicitly on the curve, $\alpha(t)$,
or, equivalently, the curve, $q(t)$.
%via $v_{*i}(T)$.
%
%If $\alpha(t)$ is a constant, $a_0$,
%so that $q(t) = a_0 \, t$,
%then after a long period,
%the error approach an equilibrium value
%%
%\begin{equation}
%  \Err_\mathrm{equil.}
%  = \Err_A = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
%  = \frac{1}{2} a_0 \Gamma
%  .
%  \label{eq:error_equil_sbin}
%\end{equation}
%
We wish to find the $q(t)$
that minimizes this error.
%
It is convenient to
vary the curve, $q(t)$ ($0 < t < T$),
under a fixed value of $q(T)$,
since it allows us to focus on
the asymptotic error, $E_A(T)$.
%
The contribution of the residual error
will be considered later
by adjusting the value of $q(T)$.
%
Note that the asymptotic error
has the form of an action of a free particle
at $\theta\bigl( q(t) \bigr)$,
with the endpoints fixed at
$\theta\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$\theta\bigl( q(T) \bigr) = 1$.
%
Thus, the velocity should be a constant
%Since the residual error
%is fixed during the process,
%the asymptotic error,
%resembling the action of a free particle,
%demands a constant velocity
%
\begin{equation}
  \frac{d}{dt} \theta\bigl( q(t) \bigr) = c
  .
\label{eq:dthetadt_const}
\end{equation}
%
Then, by using Eq. \eqref{eq:theta_def}, we have
%
\begin{equation}
  e^{ q(t) - q(T) }
  =
  c \, (t + t_0)
  =
  \frac{ t + t_0 }
       { T + t_0 }
  ,
\label{eq:expqt}
\end{equation}
%
where $t_0$ is some constant,
and the previous constant, $c$, has been determined
from the $t = T$ case
in the second step.
%
Taking the logarithm,
and differentiating both sides
with respect to $t$,
we get the inverse-time schedule
%
\begin{equation}
  \alpha(t) = \frac{ 1 }{ t + t_0 }
  ,
\label{eq:alpha_invt1}
\end{equation}
%
with Eq. \eqref{eq:alpha_invt} being
the special case of $t_0 = 0$.
%

We now determine the optimal value of $q(T)$,
which was previously held fixed.
%
This is equivalent to choosing a value of $t_0$
because $q(T) = \ln\bigl(1 + \frac{T}{t_0}\bigr)$
from the $t = 0$ case of Eq. \eqref{eq:expqt}.
%
Using Eq. \eqref{eq:alpha_invt1} in
Eqs. \eqref{eq:ER_sbin} and \eqref{eq:EA_sbin}
yields
\begin{align}
  \Err_R(T)
  &= \frac{ \Err(0) \, t_0^2 } { (T + t_0)^2 }
  ,
  &
  %\notag
  %\label{eq:ER_sbin1}
  \\
  \Err_A(T)
  &= \frac{ \Gamma \, T } { (T + t_0)^2 }
  ,
  %\label{eq:EA_sbin1}
\end{align}
%
and the total has a minimum of
\begin{equation}
  \Err(T)
  =
  \frac{ \Gamma } { T + t_0 }
  ,
  \label{eq:Emin_sbin}
\end{equation}
%
reachable at $t_0 = \Gamma /\Err(0)$.
%
%If we assume that at $t = 0$,
%the system has just completed a long simulation
%under a constant updating magnitude, $a_0$,
%then the initial error can be computed from Eq. \eqref{eq:EA_sbin}
%as
%\begin{equation}
%  \Err(0)
%  = \Gamma \int_0^\infty e^{-2\,q'} \, a_0 \, d q'
%  = \frac{1}{2} a_0 \Gamma
%  ,
%  \label{eq:error_equil_sbin}
%\end{equation}
%and
%%
%\begin{equation}
%  t_0 = \frac{ 2 } { a_0 } = \frac{1}{ \alpha(t = 0) }
%  .
%  \label{eq:t0_sbin}
%\end{equation}
%
%The minimum error can be related to
%the expected histogram fluctuation\cite{zhou2005, zhou2008}
%(cf. Appendix \ref{sec:hfluc}).

@@

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \notag
  %\label{eq:w_balance}
\end{equation}
%
We will label this eigenvector by index $k = 0$,
and
%
\begin{equation}
  \phi_{0i} = \rho_i,
\label{eq:eigenmode0}
\end{equation}
%
with $\lambda_0 = 1$.
%
Clearly, Eq. \eqref{eq:eigenmode0}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 0$,
while the orthogonality demands
%
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \delta_{k0}
  .
\label{eq:ortho0}
\end{equation}

@@

Equation \eqref{eq:vt_diffeq_mbin} shows that
in the absence of the fluctuating term, $\tilde \xi_k(t)$,
the deviation of the bias potential would tend to zero
at a rate of $\alpha(t) \, \lambda_k$.
%
Thus, the updating scheme is convergent
only if no eigenvalue, $\lambda_k$, is negative
for $k > 0$.
%
%With a set of heterogeneous eigenvalues,
%the errors of the eigenmodes are reduced at different rates.
%Thus, the optimization problem of the final error
%becomes more complicated and it often
%involves a compromise among eigenmodes.
%
The formal solution of Eq. \eqref{eq:vt_diffeq_mbin}
is given by
\begin{equation}
  \tilde v_k(T)
  =
  \tilde v_k(0) \, e^{-\lambda_k \, q(T)}
  +
  \int_0^T
    \dot{\theta}_k\bigl( q(t) \bigr) \, \tilde\xi_k(t) \, dt
  ,
  \label{eq:vk_solution}
\end{equation}
where
\begin{equation}
  \theta_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  .
  \label{eq:uk_def}
\end{equation}

@@

By Eq. \eqref{eq:parseval},
we can rewrite the error defined in Eq. \eqref{eq:error_def} as
a sum of the errors in individual modes\footnote{If
the error weighting distribution differs from $\rho_i$
in Eq. \eqref{eq:error_def},
the error would contain quadratic cross terms among different eigenmodes.
%
While the solution for the optimal schedule retains the same form,
the mass function defined in Eq. \eqref{eq:mass_func}
will include the corresponding cross terms.
}
%$\Err(T) = \sum_{k = 1}^{n - 1} \left\langle {\tilde v}_k^2(T) \right\rangle$.
%
\begin{align}
  \Err(T)
  =
  \sum_{k = 1}^{n - 1}
    \left\langle
      {\tilde v}_k^2(T)
    \right\rangle
  .
  \notag
\end{align}
%
As in Eq. \eqref{eq:error_tot},
we can decompose the error as the sum
of the residual error and the asymptotic error, with
\begin{align}
  \Err_R(T)
  &=
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) }
  ,
  \label{eq:error_res}
  \\
  \Err_A(T)
  &=
  \sum_{k=1}^{n-1}
  \int_0^T\!\!\!\int_0^T
  \!\!
  \langle
    \tilde\xi_k(t)
    \tilde\xi_k(t')
  \rangle
  \theta_k\bigl(q(t)\bigr)
  \theta_k\bigl(q(t')\bigr)
  \, dtdt'
  \notag \\
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot \theta_k^2\bigl( q(t) \bigr) \, dt
  ,
  \label{eq:error_asym}
\end{align}
%
where
%
%$\Gamma_k$ is computed from
%the autocorrelation function
%of the histogram fluctuation,
%$\bigl\langle {\tilde \xi}_k(0)
%\, {\tilde \xi}_k(t) \bigr\rangle$,
%as:
%
\begin{equation}
  \Gamma_k
  \approx
  \int_{-\infty}^\infty
  \bigl\langle
    \tilde\xi_k(t) \, \tilde\xi_k(0)
  \bigr\rangle
  \, dt
  \to
  \sum_{t = -\infty}^\infty
  \bigl\langle
    \tilde\xi_k(t) \, \tilde\xi_k(0)
  \bigr\rangle
  ,
  \label{eq:Gamma_sum}
\end{equation}
%
with the second expression being
the counterpart in the discrete-time setting.
%
Equation \eqref{eq:Gamma_sum} represents
a first approximation for modeling
a general fluctuation, ${\tilde \xi}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
%
Note that by Eq. \eqref{eq:parseval}, we have
$\sum_{k=1}^{n-1} \bigl\langle {\tilde \xi}_k(t) \, {\tilde \xi}_k(0) \bigr\rangle
= \sum_{i=1}^n \rho_i \bigl\langle \xi_{*i}(t) \, \xi_{*i}(0) \bigr\rangle,$
and summing $t$ from $-\infty$ to $\infty$ yields,
\begin{equation}
  \sum_{k=1}^{n-1} \Gamma_k = \sum_{i=1}^n \rho_i \, G_i = \Gamma
  .
  \label{eq:Gammak_sum}
\end{equation}
This shows the equivalence of
Eq. \eqref{eq:EA_sbin} and \eqref{eq:error_asym}
in the case of the single-bin updating scheme, with $\lambda_k \equiv 1$.
%In Appendix \ref{sec:Gamma_measure},
%we discuss a way of estimating $\Gamma_k$
%from an FDS simulation under a constant magnitude.
%

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
To model the initial error for Eq. \eqref{eq:error_res},
we will assume that at $t = 0$
the system has undergone a preliminary FDS simulation
under a constant updating magnitude, $a_0$,
for $T_0$ steps.
%
If the initial bias potential
before the preliminary run is zero,
we have, from Eq. \eqref{eq:vk_solution} [with $T\to T_0, q(t) \to a_0 \, t$],
%
\begin{align}
  \left\langle {\tilde v}_k(0) \right\rangle
  &=
  \tilde u_k \, e^{-\lambda_k \, a_0 \, T_0}
  ,
  \label{eq:xt_eql}
  \\
  %\operatorname{var} \tilde v_k(0)
  %&=
  %\frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
  %\left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
  %,
  %\label{eq:xtvar}
  %\\
  \left\langle \tilde v_k^2(0) \right\rangle
  &=
  %\operatorname{var} \tilde v_k(0) + \zeta_k^2
  %=
  \frac 1 2 \, \Gamma_k \, \lambda_k \, a_0
  +
  \epsilon_k \, e^{-2 \, \lambda_k \, a_0 \, T_0}
  ,
  \label{eq:xt2_eql}
\end{align}
%
where
$\epsilon_k \equiv \left(\tilde u_k^2 - \frac{1}{2} \Gamma_k \, \lambda_k \, a_0\right) \, e^{-2\,\lambda_k a_0 T_0}$.

@@

The results for the $L = 64$ case
after $10^5 L^2$ steps are shown
in Fig. \ref{fig:pt_hist}.
%
The Gaussian histograms were evenly spaced with similar widths as intended,
%and the overall energy histogram was roughly flat,
as shown in Fig. \ref{fig:pt_hist}(a).
%
We computed the density of states by WHAM,
and found the critical temperature to be $T_c = 0.70163$,
in agreement with a previous study\cite{wang2001pre}.
%
Since umbrellas were relatively narrow,
the logarithmic density of states, $\ln g(E)$,
was roughly captured
by $\hat c_0^{(s)}$ as a function of $E_c^{(s)}$,
as shown Fig. \ref{fig:pt_hist}(b).
%
As discussed in Sec. \ref{sec:age},
$c_1^{(s)}/\sigma_E$ roughly represents
the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
and it manifested
the backbending behavior\cite{kim2006, *kim2007, kim2010}
typical in a first-order phase transition,
as shown in Fig. \ref{fig:pt_hist}(c).
%
The coefficient, $\sqrt 2 \, c_2^{(s)}$,
was indeed above $1.0$
in the unstable region,
where $\beta(E)$ is increasing or, equivalently,
$\ln g(E)$ is locally convex,
as shown in Fig. \ref{fig:pt_hist}(d).
